MSB1015 Project

### 2024-10-18 Code/MSB1015-Project.Rmd
Update MSB1015-Project.Rmd
Added a chunk to find the matching train and test splits with the best model

# Plot where those missing values were found
ggplot(missing_counts, aes(x = reorder(column, missing_count), y = missing_count)) +
  geom_bar(stat = "identity", fill = "royalblue") +
  coord_flip() +
  labs(x = "Column", y = "Missing Count", title = "Columns with More Than 10 Missing Values")

@@

@@

@@

  # Return the best model and the list of models for each repetition
  return(list(best_model_overall = best_model_overall, store_models_repeated = store_models_repeated, train = train, test = test))
}
```

# 2.2 Calling the random forest function
```{r perform random forest, echo=FALSE}
# Categorical results
results_categorical <- random_forest(categorical_data, ntree = 10, nrun = 2)

# Genetic results
results_genetic <- random_forest(genetic_data, ntree = 10, nrun = 2)

# Complete results
results_complete <- random_forest(complete_data, ntree = 10, nrun = 2)
```

@@

@@

# Get predicted probabilities for the positive class ("will_die_within_10_years")
pred_probs <- predict(best_class_data$best_model_overall, newdata = best_class_data$train, type = "prob") # Predict probabilities
positive_class_probs <- pred_probs[, "will_die_within_10_years"] # Extract probabilities for "will_die_within_10_years"

@@

# 5. Testing on the best model
```{r testing, echo=FALSE}
# Get predicted probabilities for the positive class ("will_die_within_10_years")
pred_probs <- predict(best_class_data$best_model_overall, newdata = best_class_data$test, type = "prob")
positive_class_probs <- pred_probs[, "will_die_within_10_years"] # Select probabilities for "will_die_within_10_years"

# Generate the ROC curve for the test set
roc_curve <- roc(best_class_data$test$prognosis, positive_class_probs, levels = levels(best_class_data$test$prognosis), direction = ">")

# Create Precision-Recall curve for the test set
pr_curve <- pr.curve(
  scores.class0 = positive_class_probs, # Probabilities for the positive class
  weights.class0 = as.numeric(best_class_data$test$prognosis == "will_die_within_10_years"), # Binary vector for positive class
  curve = TRUE # Return curve for plotting
)

@@

# Ensure predicted classes have the correct factor levels
pred_best_threshold <- factor(pred_best_threshold, levels = levels(best_class_data$test$prognosis))

# Calculate accuracy on the test set using the best threshold
pred_table <- table(pred_best_threshold, best_class_data$test$prognosis) # Create confusion matrix table
dimnames(pred_table) <- list( # Assign dimension names to match factor levels
  c("will_die_within_10_years", "will_die_within_5_years"),
  c("will_die_within_10_years", "will_die_within_5_years")
)
test_confusion <- confusionMatrix(pred_table) # Compute confusion matrix
test_confusion

>>>>

# Plot where those missing values were found
ggplot(missing_counts, aes(x = reorder(column, missing_count), y = missing_count)) +
  geom_bar(stat = "identity", fill = "royalblue") +
  coord_flip() +
  labs(x = "Column", y = "Missing Count", title = "Columns with More Than 10 Missing Values") +
  theme(
    panel.background = element_rect(fill = "white", color = "white"),
    plot.background = element_rect(fill = "white", color = "white"),
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_line(color = "gray90"),
    axis.text.y = element_text(size = 14),  
    axis.text.x = element_text(size = 12),  
    axis.title = element_text(size = 14),
    plot.title = element_text(size = 15, face = "bold")
  )

@@

  # Initialize lists to store all train and test splits
  train_splits <- vector("list", nrun)
  test_splits <- vector("list", nrun)

@@

    # Store the train and test sets for this run
    train_splits[[repeat_run]] <- train
    test_splits[[repeat_run]] <- test

@@

  # Return the best model and the list of models for each repetition
  return(list(best_model_overall = best_model_overall, 
              store_models_repeated = store_models_repeated, 
              train_splits = train_splits, 
              test_splits = test_splits
              ))
}
```

# 2.2 Calling the random forest function
```{r perform random forest, echo=FALSE}
# Categorical results
results_categorical <- random_forest(categorical_data, ntree = 1000, nrun = 5)

# Genetic results
results_genetic <- random_forest(genetic_data, ntree = 10, nrun = 1)

# Complete results
results_complete <- random_forest(complete_data, ntree = 10, nrun = 1)
```

@@

# 3.3 Finding the matching splits for the best model
```{r finding matching splits}
# Get the AUC value of the best model
best_auc <- best_class_data$best_model_overall$roc$auc

# Initialize variables to store the matching run and model
matching_run <- NULL
matching_model <- NULL
matching_index <- NULL

# Loop over all runs in store_models_repeated
for (run_index in seq_along(best_class_data$store_models_repeated)) {
  # Get the current run's list of models
  current_run_models <- best_class_data$store_models_repeated[[run_index]]

  # Loop over all models in the current run and compare their AUC values
  for (model_index in seq_along(current_run_models)) {
    model_name <- names(current_run_models)[model_index]
    model_auc <- current_run_models[[model_name]]$roc$auc

    # Check if the AUC matches the best model's AUC
    if (model_auc == best_auc) {
      matching_run <- run_index
      matching_model <- model_name
      matching_index <- model_index
      break
    }
  }

  # Break out of the outer loop if a match is found
  if (!is.null(matching_run)) {
    break
  }
}

# If a match is found, print the run, model name, and retrieve the corresponding train/test splits
if (!is.null(matching_run) && !is.null(matching_model)) {
  print(paste("Matching model found in run:", matching_run, "model:", matching_model))

  # Retrieve the corresponding train and test splits
  matching_train_split <- best_class_data$train_splits[[matching_run]]
  matching_test_split <- best_class_data$test_splits[[matching_run]]

  # Print or use the splits as needed
  print("Found Matching Splits")

} else {
  print("No matching model found")
}

```

@@

# Get predicted probabilities for the positive class ("will_die_within_10_years")
pred_probs <- predict(best_class_data$best_model_overall, newdata = matching_train_split, type = "prob") # Predict probabilities
positive_class_probs <- pred_probs[, "will_die_within_10_years"] # Extract probabilities for "will_die_within_10_years"

@@

# 5. Testing on the best model
```{r testing, echo=FALSE}
# Get predicted probabilities for the positive class ("will_die_within_10_years")
pred_probs <- predict(best_class_data$best_model_overall, newdata = matching_test_split, type = "prob")
positive_class_probs <- pred_probs[, "will_die_within_10_years"] # Select probabilities for "will_die_within_10_years"

# Generate the ROC curve for the test set
roc_curve <- roc(matching_test_split$prognosis, positive_class_probs, levels = levels(matching_test_split$prognosis), direction = ">")

# Create Precision-Recall curve for the test set
pr_curve <- pr.curve(
  scores.class0 = positive_class_probs, # Probabilities for the positive class
  weights.class0 = as.numeric(matching_test_split$prognosis == "will_die_within_10_years"), # Binary vector for positive class
  curve = TRUE # Return curve for plotting
)

@@

# Ensure predicted classes have the correct factor levels
pred_best_threshold <- factor(pred_best_threshold, levels = levels(matching_test_split$prognosis))

# Calculate accuracy on the test set using the best threshold
pred_table <- table(pred_best_threshold, matching_test_split$prognosis) # Create confusion matrix table
dimnames(pred_table) <- list( # Assign dimension names to match factor levels
  c("will_die_within_10_years", "will_die_within_5_years"),
  c("will_die_within_10_years", "will_die_within_5_years")
)
test_confusion <- confusionMatrix(pred_table) # Compute confusion matrix
test_confusion
